Chronic conditions and antibiotic consumption among the non-elderly

### 2017-03-24 chronic-conditions/young-people/cc-young.Rmd
young chronic

---
title: "Chronic conditions and increased antibiotic consumption"
author: "Scott Olesen"
date: "2/27/2017"
output: html_document
---

@@

Just summing up the chronic conditions doesn't seem like a very elegant way to
summarize people's health. Choosing which chronic conditions to consider could
be a matter of expert opinion. It's also worth asking which ones are
interesting, from the data's viewpoint.

@@

I'll use the Medicare 2011 data with normal filters (age 65 or above, only beneficiaries on Medicare for 12 months of the year).

@@

```{r load_data}
denom = read_tsv('../bene_2011.tsv') %>%
  filter(age >= 65) %>%
  mutate(is_female=sex=='female') %>%
  select(bene_id, age, is_female)
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
usage = read_tsv('../pde_2011.tsv') %>%
  count(bene_id) %>% rename(n_claims=n)

@@

usage_summary = denom %>%
  left_join(chronic_conditions, by='bene_id') %>%
  left_join(usage, by='bene_id') %>%
  mutate(n_claims=na_to_0(n_claims))

condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
```

@@

res = lapply(conditions, function(cond) summarize_condition(usage_summary, cond)) %>%
  bind_rows() %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name)
```

```{r multivariate_poisson}
multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>% formula

multivariate_res = usage_summary %>%
  select(n_claims, age, is_female, AMI:HYPOTH) %>%
  glm(multivariate_formula, data=., family='poisson') %>%
  tidy
```

@@

lasso_x = usage_summary %>%
  select(age, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage_summary$n_claims

@@

## Summary

**XXX**

@@

Presumably many of the conditions are correlated. I computed tetrachoric correlations, which is a method designed to deal with correlations of dichotomous variables.

@@

Thankfully, Alzheimer's correlates strongly with Alzheimer's and dementia, and prostatic hyperplasia and prostate cancer correlate strongly negatively with breast cancer and endometrial cancer. Unfortunately, the negative correlations make everything else harder to see, so here's the same thing but without the male conditions.

@@

COPD correlates strongly with asthma and lung cancer, but lung cancer and asthma do not correlate strongly with one another. There is a block of correlations focused around hypertension that includes ischemic heart disease, heart failure, anemia, chronic kidney disease, diabetes, hypertension, hyperlipidemia, and, importantly, COPD.

@@

Almost all the conditions are significantly associated with greater consumption. Here I show the results of the Poisson regression: the column `estimate` is the estimate for the parameters associated with each condition, and `p.value` measures whether the data are consistent with that parameter being zero. The column `pct.increase` shows that percent increase in number of claims derived from that estimate.

@@

As a verification, I'll look at the distribution of number of claims among patients who have COPD or not. The mean for patients with COPD is more than double than that for patients without, which makes the age-adjusted 2.23 fold increase seem reasonable.

```{r copd}
usage_summary %>%
  group_by(COPD) %>%
  do(summary(.$n_claims) %>% tidy) %>%
  kable
```

### Multivariate

```{r show_multivariate, cache=FALSE}
# Here I need to do some fancy footwork to get the condition names in
# The terms look like (Intercept), age, is_female, AMITRUE, COPDTRUE, etc.
# So chop off TRUE where it is, then match with the known names, and
# keep the (Intercept), etc. if it didn't match

multivariate_res %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, pct.increase, estimate, p.value) %>%
  arrange(desc(abs(estimate))) %>%
  kable
```

@@

Not surprisingly, COPD tops the list of chronic conditions associated with increased consumption. Interestingly, asthma does provide explanatory power beyond COPD, coming in second. Depression and anemia are next. Kidney disease has a middling effect, while age and diabetes have almost no effect!

The lasso results seem to be the nicest: they sum up the major findings very well but seem to avoid some of the weird noise (e.g., enlarged prostate as the #2 coefficient in the multivariate Poisson model, lung cancer as the #3 coefficient in the univariate models).

>>>>

---
title: "Chronic conditions and antibiotic consumption among the non-elderly"
author: "Scott Olesen"
date: "2/27/2017"
output: html_document
---

@@

We see that chronic conditions, not age, control consumption among the elderly.
Is the same true among the non-elderly? The Medicare people are going to be
pretty sick.

@@

I'll use the Medicare 2011 data with different filters: age below 65, on
Medicare for 12 months of the year).

@@

```{r load_data}
bene = read_tsv('../../bene_2011.tsv') %>%
  filter(age < 65) %>%
  mutate(is_female=sex=='female') %>%
  select(bene_id, age, is_female)
chronic_conditions = read_tsv('../cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
pde_summary = read_tsv('../../pde_2011.tsv') %>%
  count(bene_id) %>% rename(n_claims=n)

@@

usage = bene %>%
  left_join(chronic_conditions, by='bene_id') %>%
  left_join(pde_summary, by='bene_id') %>%
  mutate(n_claims=na_to_0(n_claims))

condition_names = read_tsv('../names.txt', col_names=c('condition', 'condition_name'))
```

@@

res = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
  bind_rows() %>%
  left_join(condition_names, by='condition') %>%
  select(-condition) %>%
  rename(condition=condition_name)
```

@@

lasso_x = usage %>%
  select(age, is_female, AMI:HYPOTH) %>%
  as.matrix
lasso_y = usage$n_claims

@@

@@

@@

Without male conditions (prostate cancer, enlarged prostate):

@@

@@

Here I show the results of the Poisson regression: the column
`estimate` is the estimate for the parameters associated with each condition,
and `p.value` measures whether the data are consistent with that parameter
being zero. The column `pct.increase` shows that percent increase in number of
claims derived from that estimate.

@@

@@
